---
title: "seer: R package for featue-based forecast model selection"
author: 
- Thiyanga S Talagala \newline
- Rob J Hyndman \newline
- George Athanasopoulos
institute: |
        | Monash University, Australia
date: "UseR, 2018"
output: 
  beamer_presentation:
     theme: "Warsaw"
     includes:
        in_header: header_pagenrs.tex
---

```{r initial, echo = FALSE, cache = TRUE, include = FALSE}
library(knitr)
library(pander)
library(seer)
library(Mcomp)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
opts_chunk$set(fig.lp = '')
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.path = "figure/",
  cache.path = "cache/",
  cache = TRUE,
  fig.show = "hold",
  external = TRUE,
  dev = "png"
)

```

```{r set-options, echo=FALSE, cache=TRUE}
options(width = 50)
```


# Large collections of time series 

  \centerline{\includegraphics[width=\textwidth,height=7.5cm,keepaspectratio]{figures/example.jpg}}

- Forecasting demand for thousands of products across multiple
warehouses.


# Time series features

\begin{block}{Objective}
Develop a framework that automates the selection of the most appropriate forecasting method for a given time series by using an array of \textcolor{red}{features} computed from the time series.
\end{block}

\pause
\begin{itemize}[<+->]

\item \textbf{Basic idea:}
  
  Transform a given time series $y=\{y_1, y_2, \cdots, y_n\}$ to a feature vector $F = (f_1(y), f_2(y), \cdots, f_p(y))'$. 

\item Examples for time series features

\begin{itemize}[<+->]
\vfill\item strength of trend
\vfill\item strength of seasonality
\vfill\item lag-1 autocorrelation
\vfill\item spectral entropy

\end{itemize}

\end{itemize}


# Time series features



\begin{columns}
\begin{column}{0.48\textwidth}
   \begin{itemize}
   \item length
    \item strength of seasonality
    \item strength of trend
    \item linearity
    \item curvature
    \item spikiness
    \item stability
    \item lumpiness
    \item first ACF value of remainder series
    \item parameter estimates of Holt's linear trend method
    \end{itemize} 
\end{column}
\begin{column}{0.48\textwidth}
   \begin{itemize}
    \item spectral entropy
    \item Hurst exponent
    \item nonlinearity
    \item parameter estimates of Holt-Winters' additive method
    \item unit root test statistics
    \item first ACF value of residual series of linear trend model 
    \item ACF and PACF based features - calculated on both the raw and differenced series
    \end{itemize} 
\end{column}
\end{columns}

# Methodology: FFORMS

**FFORMS:** **F**eature-based **FOR**ecast **M**odel **S**election

\textcolor{red}{Offline}

- A classification algorithm (the meta-learner) is trained.

\textcolor{red}{Online}

- Calculate the features of a time series and use the pre-trained classifier to identify the best forecasting method.


# FFORMS: population

 \centerline{\includegraphics[width=\paperwidth]{figures/fw1.png}}

# FFORMS: observed sample

 \centerline{\includegraphics[width=\paperwidth]{figures/fw2.png}}
 
# FFORMS: simulated time series

 \centerline{\includegraphics[width=\paperwidth]{figures/fw3.png}}
 
# FFORMS: reference set
 
 \centerline{\includegraphics[width=\paperwidth]{figures/fw4.png}}

# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw5.png}}

# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw6.png}}

# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw7.png}}

# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw8.png}}


# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw9.png}}

# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw10.png}}

# FFORMS: Random-forest classifier

\centerline{\includegraphics[width=\paperwidth]{figures/fw11.png}}


# FFORMS: "online" part of the algorithm

\centerline{\includegraphics[width=\paperwidth]{figures/fw12.png}}


# FFORMS: "online" part of the algorithm

\centerline{\includegraphics[width=\paperwidth]{figures/fw13.png}}

# FFORMS: "online" part of the algorithm

\centerline{\includegraphics[width=\paperwidth]{figures/fw14.png}}


# Discussion and Conclusions

\begin{itemize}[<+->]

\vfill\item FFORMS: framework for forecast model selection using meta-learning based on time series features.

\vfill\item FFORMS algorithm uses the knowledge of the past performance of candidate forecast models on a collection of time series in order to identify the best forecasting method for a new series. 

\vfill\item For real-time forecasting, our framework involves only the calculation of features, the selection of a forecast method based on the FFORMS random forest classifier, and the calculation of the forecasts from the chosen model.

\vfill\item We have also introduced a simple set of time series features that are useful in identifying the "best" forecast method for a given time series.

\end{itemize}


# R package: seer

 \vspace{0.15cm}
\centerline{\includegraphics[width=\textwidth,height=3.5cm,keepaspectratio]{figures/seer.png}}
\vspace{0.3cm}
 \centerline{available at: \textcolor{blue}{\underline{https://github.com/thiyangt/seer}}}

## Installation

```r
devtools::install_github("thiyangt/seer")
library(seer)
```
\pause

##
paper: https://robjhyndman.com/publications/fforms/

##

Email: thiyanga.talagala@monash.edu

